---
title: "Assignment 3-GGplot"
author: "Wagner Quirós"
date: "May 18, 2017"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
setwd("/Users/wagnerquiros/Desktop/UCSB/Courses/Spring 2017/Informatics/Assignments_ESM262/esm262/docs/wk04_ggplot")
thinshallow = read.table("data/resthin.shallow.txt", header=T)
thindeep = read.table("data/resthin.deep.txt", header=T)
clim = read.table("data/sierraclim.txt", header = T)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
```


"Part 1. Clim Figure with Temperature and Rain"

```{r 1, fig.align='center'}
#Monthly average temperature from tmax and tmin
clim$avg.temp <- (clim$tmin + clim$tmax)/2

#Plot Figure 1.
Figure_1 = ggplot(clim, aes(y=avg.temp,x=year))+
  stat_summary(fun.y="mean", geom="point")+
  stat_smooth()+
  theme_bw()+
  labs(x="", y="Average Temperature (Celsius)")+
  theme(axis.text.x=element_blank())+
  theme(axis.ticks.x=element_blank())
Figure_1

#Plot 2
Figure_2 = ggplot(clim, aes(y=rain, year))+
  stat_summary(fun.y="mean", geom="point")+
  stat_smooth()+theme_bw()+
  labs(x="Year", y="Average Rain (Inches)")
Figure_2

#Get both figures together.
grid.arrange(Figure_1, Figure_2)
```

"Part 2. Thindeep Figure "


**BIOMASS OVER TIME BEHAVIOR**
```{r color}
View(thindeep)
p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))
p1
# but boxplots include all thinning intensities - maybe we want to separate those out

p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc, col=as.factor(thin)))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))

# or we could base the filling of the boxplots on thinning intensity

p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc, fill=as.factor(thin)))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))
p1

```




This is a good opporutnity to play witht e legend a little, we can use *theme* again to control formatting
and *scale_color_discrete* to control content
There are many other ways to change color maps - see

* *scale_colour_hue*
* *scale_fill_hue*
* *scale_color_brewer*
* *scale_color_brewer*

Brewer is really cool and comes up with nice options for you  and gives
you a menu of color scheme for sequential vs divergin vs qualitative data and 
has optins that are printer friendly, color blind friendly 
-to check it out - 

* [colors that work](http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)
* [implemented R color palettes](https://www.r-bloggers.com/choosing-colour-palettes-part-ii-educated-choices/)
* [more R and colors](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)

or for continues
* *scale_color_continuous*

```{r legendsandstuff}

# assuming qualitative
 p1 + scale_fill_brewer(type="qual", palette="Dark2")
 p1 + scale_fill_brewer(type="qual", palette="Set3")
 
# but actually sequential so can using diverging palettes - diverse from center
 p1 + scale_fill_brewer(type="div", palette="PiYG")
 
 # but actually sequential so can using diverging palettes - diverse from center
  p1 + scale_fill_brewer(type="seq", palette="BuGn")

 # and can add names
   lnms = c("None","Low","Med","High")
   p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms)
   
# using themes to change legend format - position. CHANGE LEGEND.
    p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
      theme(legend.position="bottom")
    
     p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9)) #locates the quadrant of the legend.
     
  p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9), legend.background=element_rect(fill="seashell2"))    
        
```

We can also use ggplot to automatically summarize graph and added to our plot
Lets say we also want to show the means as lines


```{r summaries}

# what if we just want means
# there are multiple ways to do this but 

p1=ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(thin)))+stat_summary(fun.y="mean", geom="line")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14))
p1 + scale_color_brewer(type="div", palette="Spectral", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9), legend.background=element_rect(fill="seashell2"))    

# we also might want to separate out our two different types of thinning - indicated by shared
p1=ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(thin), type=shared))+stat_summary(fun.y="mean", geom="line", aes(linetype=shared))
p1 = p1 +theme(axis.text= element_text(face="bold", size=14))
p1 + scale_color_brewer(type="div", palette="Spectral", name="Thin Intensity", labels=lnms) + 
  scale_linetype(name="Thin Type", labels=c("Spaced","Clustered"))

#averaging by year and then fitting a curve        
p1 = ggplot(clim, aes(y=tmax,x=year))+stat_summary(fun.y="mean", geom="point")+stat_smooth()
p1
# we can also combined bar plots and line summaries (and deal with different axis)
# we need to transform to get similar scales

p1 = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum") + ggtitle("Rain")
p1
p2 = ggplot(clim) + geom_line(aes(x=year, y=tmax), stat="summary", fun.y="mean") + ggtitle("Temp")
p2
grid.arrange(p1,p2)
p2 + geom_line(aes(x=year, y=tmin), stat="summary", fun.y="mean")

# estimate scaling so that we can pput the values in the same range.
scl = 3000/15
# add some labels

p = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum",  fill="cyan") + 
  geom_line(aes(x=year, y=tmax*scl), stat="summary", fun.y="mean", col="red") 
p 
p=p+ scale_y_continuous(sec.axis = sec_axis(~./scl, name=expression(paste("Maximum Temperature ",C**degree)))) + 
  labs(x="Year", y="Rainfall (mm/yr)")
p


# annotation. Put text on the plot.!! COOL.!!
p = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum",  fill="cyan") + 
  geom_line(aes(x=year, y=tmax*scl), stat="summary", fun.y="mean", col="red") 
p=p+ scale_y_continuous(sec.axis = sec_axis(~./scl, name=expression(paste("Maximum Temperature ",C**degree)))) + 
  labs(x="Year", y="Rainfall (mm/yr)")
p = p+ annotate("text", x=1990, y=500, label="Precip", colour="blue", size=6, hjust=0.5)
p = p+ annotate("text", x=1940, y=2700, label="Temperature", colour="red", size=6, hjust=0)
p

```

Now lets try *facets*  which allow you to separate things into different graphs
lets say we wanted a separate graph for each thinning intensity so  we could see how decade impacted recovery

```{r facet}
# first lets color by decade/scenario. Facet creates multiple graphs base on a dimension!!!Secen is the DECADE.
p2 = ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(scen)))+ stat_summary(fun.y="mean",geom="line", aes(col=as.factor(scen)))
p2
# now lets separate out the thinning intensities
p2= p2+facet_wrap(~as.factor(thin))
p2

# make prettier
p2 = ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(scen)))+ stat_summary(fun.y="mean",geom="line", size=2, aes(col=as.factor(scen)))
p2= p2+facet_wrap(~as.factor(thin)) 
p2 + scale_color_brewer(type="qual", palette="Set3", name="Decade") 
p2
```

