---
title: "Assignment_2.Informatics."
author: "Wagner Quiros Pereira"
date: "5/9/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ESM ASSIGNMENT 2.

##Part1. Data Ingest.

**1. Reading the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble and keeping only specific columns**

```{r 1, fig.align='center'}
suppressPackageStartupMessages({
  library(tidyverse)
  library(dplyr)
  library(magrittr)
  library(knitr)
})

gaz_raw<- read_delim("CA_Features_20170401.zip", delim = "|")
View(gaz_raw)
gaz <- gaz_raw %>% 
  select (-COUNTY_NUMERIC, STATE_NUMERIC, PRIMARY_LAT_DMS, PRIM_LONG_DMS, SOURCE_LAT_DMS, SOURCE_LONG_DMS, ELEV_IN_FT) %>% 
  mutate(DATE_CREATED = as.Date(DATE_CREATED, format= "%m/%d/%Y"), 
         DATE_EDITED = as.Date(DATE_EDITED, format= "%m/%d/%Y" ))
as_tibble(gaz_raw)
```

**2. Converting the gaz columns to the appropriate type, and any placeholders for unknown data to NA**

```{r 2, fig.align='center'}
gaz$PRIM_LAT_DEC <- ifelse(gaz$PRIM_LAT_DEC == 0, NA, gaz$PRIM_LAT_DEC)
gaz$PRIM_LONG_DEC <- ifelse(gaz$PRIM_LONG_DEC == 0, NA, gaz$PRIM_LONG_DEC)
gaz$MAP_NAME <- ifelse(gaz$MAP_NAME =='Unknown', NA, gaz$MAP_NAME)
View(gaz)
```

**3. Selecting main observations**

```{r 3, fig.align='center'}
gaz <- gaz %>%
  filter(!(PRIM_LAT_DEC == 'NA' | PRIM_LONG_DEC == 'NA')) %>% 
  filter(STATE_ALPHA == 'CA')
View(gaz)
```

**4. Saving clean up doc with delim |**

```{r 4, fig.align='center'}
write_delim(gaz, "/Users/wagnerquiros/Desktop/UCSB/Courses/Spring 2017/Informatics/newprojects/gaz.csv", delim = "|")
View(gaz)
```


##Part2. Analyze.
**1.What is the most-frequently-occuring feature name ?**

```{r 5, fig.align='center'}
Common_feature_name <- count(gaz, FEATURE_NAME)
Common_feature_name[which.max(Common_feature_name$n), 1]
```

**The most frequestly ocurring feature name is the church of Christ*

**2.What is the most-frequently-occuring feature class ?**


```{r 6, fig.align='center'}
Common_feature_class <- count(gaz, FEATURE_CLASS)
Common_feature_class[which.min(Common_feature_class$n), 1]
```

**The most frequestly ocurring feature class is Isthmus*

**3.What is the approximate center point of each county CPC ?**

```{r 7, fig.align='center'}
c_county_point <- gaz %>%
 group_by(COUNTY_NAME) %>%
  summarise(latmin = min(PRIM_LAT_DEC, na.rm= TRUE),
            latmax = max(PRIM_LAT_DEC, na.rm= TRUE),
            longmin = min(PRIM_LONG_DEC, na.rm= TRUE),
            longmax = max(PRIM_LONG_DEC, na.rm= TRUE)) %>% 
  mutate(lat_center = (latmin + latmax) /2)%>% 
  mutate(long_center = (longmin + longmax)/2) %>% 
  select(County = COUNTY_NAME, Latitud = lat_center, 
         Longitude= long_center)
View(c_county_point)
Table1 <- table(c_county_point)
View(Table1)
```


**4.What are the fractions of the total number of features in each county that are natural? man-made?**

```{r 8, fig.align='center'}
natural <- c("Valley", "Spring" ,"Stream", "Gut", "Canal", "Cape", "Lake", "Summit", "Plain", "Woods", "Gap",   "Ridge","Bay",  "Beach",   "Channel", "Slope", "Island",  "Basin", "Cliff", "Swamp","Area" , "Bend",  "Range", "Falls",  "Forest", "Lava", "Rapids", "Crater", "Glacier", "Oilfield","Sea", "Arroyo", "Isthmus")

Feature_category <- tibble(FEATURE_CLASS= unique(gaz$FEATURE_CLASS)) %>% 
  mutate(category = ifelse(FEATURE_CLASS %in% natural == "TRUE", 1, 0)) 

#0=manmade, 1= natural
  
gaz_complete <- inner_join(gaz, Feature_category, by = "FEATURE_CLASS")

Features_type <- gaz_complete%>% 
  group_by(COUNTY_NAME) %>% 
  summarise(total = n(), natural = sum(category)) %>% 
  mutate(fraction_natural = natural/total) %>% 
  mutate(fraction_manmade = 1 - fraction_natural) %>% 
  select(County= COUNTY_NAME, Natural= fraction_natural, Man_Made= fraction_manmade)
```


```{r 9, fig.align='center'}
kable(c_county_point)
```

```{r 10, fig.align='center'}
kable(Features_type)
```
