---
title: "Assignment 3-GGplot"
author: "Wagner Quirós"
date: "May 18, 2017"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
setwd("/Users/wagnerquiros/Desktop/UCSB/Courses/Spring 2017/Informatics/Assignments_ESM262/esm262/docs/wk04_ggplot")
thinshallow = read.table("data/resthin.shallow.txt", header=T)
thindeep = read.table("data/resthin.deep.txt", header=T)
clim = read.table("data/sierraclim.txt", header = T)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
```


"Part 1. Clim Figure with Temperature and Rain"
```{r 1, fig.align='center'}
#Monthly average temperature from tmax and tmin
clim$avg.temp <- (clim$tmin + clim$tmax)/2

#Plot Figure 1.
Figure_1 = ggplot(clim, aes(y=avg.temp,x=year))+
  stat_summary(fun.y="mean", geom="point")+
  stat_smooth()+
  theme_bw()+
  labs(x="", y="Average Temperature (Celsius)")+
  theme(axis.text.x=element_blank())+
  theme(axis.ticks.x=element_blank())
Figure_1

#Plot 2
Figure_2 = ggplot(clim, aes(y=rain, year))+
  stat_summary(fun.y="mean", geom="point")+
  stat_smooth()+theme_bw()+
  labs(x="Year", y="Average Rain (Inches)")
Figure_2

#Get both figures together.
grid.arrange(Figure_1, Figure_2)
```













```{r basic example}
# basic plot
ggplot(clim, aes(y=tmax, x=month))+geom_point()

# would look bettter as a boxplot. as.factor we are saying you can group things and fixes problems in ggplot. Interpretes data as a category instead of a number. To group something by category is easier using factors.
ggplot(clim, aes(y=tmax, x=as.factor(month)))+geom_boxplot()

# save the basic plot so we can add to it
p=ggplot(clim, aes(y=tmax, x=as.factor(month)))+geom_boxplot()

p+labs(y="Maximum Temperature", x="Month") #Just adding, nit yet saving changes.
p = p + labs(y="Maximum Temperature", x="Month")
p

#Exploring lines with points and. Line 241.
Figure_1 = ggplot(clim, aes(y=tmax,x=year))+stat_summary(fun.y="mean", geom="point")+stat_smooth()
p1

```



Thre are many ways to do this (see labs and titles) but **themes** give you alot of control on your plot elements. Was to change the way the plot looks. Themes allow to change characteristics of those: Italic, bold, size, etc.!

* plot.title
* axis.title
* axis.title.x
* axis.title.y
* axis.text
* axis.text.y

I'll also show you the use of **expression** to write math symbols as labs

More on all the options that can be found with [**themes**](http://ggplot2.tidyverse.org/reference/theme.html)

```{r theme examples}
# themes are used to control character of the plot
pclr =ggplot(clim, aes(y=tmax, x=as.factor(month)))+geom_boxplot()
pclr = pclr + labs(y="Maximum Temperature", x="Month")
pclr = pclr+theme(axis.text= element_text(face="bold", colour="red", size=14))
pclr = pclr +  geom_hline(yintercept=0, col="yellow", size=4)
pclr

#lets say we want to color in the boxes - this is inside the boxplto so can't add. Col outer part and fill inside.
p = ggplot(clim, aes(y=tmax, x=as.factor(month)))+geom_boxplot(col="rosybrown3", fill="red")
p
# useful to see color options. Expression allows to add the celcious sign.
colors()

p = p + labs(y= expression(paste("Maximum Temperature ",C**degree)), x="Month") 
p= p+theme(axis.text= element_text(face="bold", colour="yellow3", size=14), 
           axis.title=element_text(face="italic", size=14))
p 
p = p+  ggtitle("Monthly Temperature in Sierras")
p


# notice how themes keep inheriting - so you can call theme multiple times
p + coord_flip() 
p=p + coord_flip() + theme(plot.margin=unit(c(1,3,1,3),"cm")) # top right bottom left
p
p = p + theme(plot.title=element_text(size=16, hjust=0.5), plot.background=element_rect(fill="peachpuff1")) 
p

# to display more than one plot (a "matrix" or grid of plots)

grid.arrange(p,pclr)
# or to control to make by col/ BOTH PLOTS TOGETHER 
grid.arrange(p, pclr, ncol=2) #changes into 2 columns.

# there are also built in themes
pclr = pclr+theme_bw()

pclr
```
There are many different geoms (geometries) - lets try the standard ones

```{r plot types}


# scatter plot
p2=ggplot(clim, aes(x=tmax, y=tmin))+geom_point(col="blue", shape=9, size=rel(4)) #ANother option is on shape you can use "*", adding symbol in quotes. (4), Making the value 4 times as big.
p2
  p2 = p2+labs(x="Max Temp C", y="Min Temp C") 
  p2 = p2+ ggtitle("How does daily maximim and min temp compare")
  p2 = p2 + geom_abline(intercept=0,slope=1, colour="green", size=4)
p2

#density plot
p3=ggplot(clim, aes(x=rain))+geom_density()
p3
p3 = ggplot(subset(clim, clim$rain > 0), aes(x=rain))+geom_density(fill="blue")
#p4=ggplot(clim, aes(x=date,y=rain))+geom_line()+ggtitle("Line Graph"). I also subsetted the days where there was rain over 0.
p3
#fix issue with date
clim$date = mdy(paste(clim$month, clim$day, clim$year, sep="/"))
p4=ggplot(clim, aes(x=date,y=rain))+geom_line()+ggtitle("Line Graph")
p4
grid.arrange(p,p2,p3,p4) # Here we arrange plots in a column with ncol.
```



INCREBIBLE USEFULL. One of the most useful things about GGPLOT is that it makes it easy to visualize your data in ways that highlight different attributes - this can help you to see multiple dimensions at once
There are multiple ways to do this

* color/size/shape (within plots)
* facets  (between plots)

Lets start with color
We can use a more interesting data set *thindeep*

This dataset is results of a simulation experiment that looks at the impacts of different levels of forest
thinning on water and carbon fluxes for a site in the California Sierra

*  The site/plot is simulated for 10 years following the thinning experiment. 
*  The thinning experiment is repeated for different start dates within a 50 year historic climate record,
*  And for different thinning intensities 
*  And for different types of thinning that try to maximize or minimize sharing of water between trees


**BIOMASS OVER TIME BEHAVIOR**
```{r color}
View(thindeep)
range(thindeep$thin) #different thining.
range(thindeep$scen) #from the number of years of data.
# ok lets look at how biomass recovers. BIOMASS OVER TIME.
p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))
p1
# but boxplots include all thinning intensities - maybe we want to separate those out

p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc, col=as.factor(thin)))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))

# or we could base the filling of the boxplots on thinning intensity

p1=ggplot(thindeep, aes(x=as.factor(wy), y=plantc, fill=as.factor(thin)))+geom_boxplot()
p1 = p1+labs(x="Years since thinning", y="Biomass")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14), 
               plot.margin = unit(c(15,15,15,5),"pt"), axis.title = element_text(size=14))
p1

```

This is a good opporutnity to play witht e legend a little, we can use *theme* again to control formatting
and *scale_color_discrete* to control content
There are many other ways to change color maps - see

* *scale_colour_hue*
* *scale_fill_hue*
* *scale_color_brewer*
* *scale_color_brewer*

Brewer is really cool and comes up with nice options for you  and gives
you a menu of color scheme for sequential vs divergin vs qualitative data and 
has optins that are printer friendly, color blind friendly 
-to check it out - 

* [colors that work](http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)
* [implemented R color palettes](https://www.r-bloggers.com/choosing-colour-palettes-part-ii-educated-choices/)
* [more R and colors](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)

or for continues
* *scale_color_continuous*

```{r legendsandstuff}

# assuming qualitative
 p1 + scale_fill_brewer(type="qual", palette="Dark2")
 p1 + scale_fill_brewer(type="qual", palette="Set3")
 
# but actually sequential so can using diverging palettes - diverse from center
 p1 + scale_fill_brewer(type="div", palette="PiYG")
 
 # but actually sequential so can using diverging palettes - diverse from center
  p1 + scale_fill_brewer(type="seq", palette="BuGn")

 # and can add names
   lnms = c("None","Low","Med","High")
   p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms)
   
# using themes to change legend format - position. CHANGE LEGEND.
    p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
      theme(legend.position="bottom")
    
     p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9)) #locates the quadrant of the legend.
     
  p1 + scale_fill_brewer(type="seq", palette="BuGn", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9), legend.background=element_rect(fill="seashell2"))    
        
```

We can also use ggplot to automatically summarize graph and added to our plot
Lets say we also want to show the means as lines


```{r summaries}

# what if we just want means
# there are multiple ways to do this but 

p1=ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(thin)))+stat_summary(fun.y="mean", geom="line")
p1 = p1 +theme(axis.text= element_text(face="bold", size=14))
p1 + scale_color_brewer(type="div", palette="Spectral", name="Thin Intensity", labels=lnms) +
       theme(legend.position=c(0.1,0.9), legend.background=element_rect(fill="seashell2"))    

# we also might want to separate out our two different types of thinning - indicated by shared
p1=ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(thin), type=shared))+stat_summary(fun.y="mean", geom="line", aes(linetype=shared))
p1 = p1 +theme(axis.text= element_text(face="bold", size=14))
p1 + scale_color_brewer(type="div", palette="Spectral", name="Thin Intensity", labels=lnms) + 
  scale_linetype(name="Thin Type", labels=c("Spaced","Clustered"))

#averaging by year and then fitting a curve        
p1 = ggplot(clim, aes(y=tmax,x=year))+stat_summary(fun.y="mean", geom="point")+stat_smooth()
p1
# we can also combined bar plots and line summaries (and deal with different axis)
# we need to transform to get similar scales

p1 = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum") + ggtitle("Rain")
p1
p2 = ggplot(clim) + geom_line(aes(x=year, y=tmax), stat="summary", fun.y="mean") + ggtitle("Temp")
p2
grid.arrange(p1,p2)
p2 + geom_line(aes(x=year, y=tmin), stat="summary", fun.y="mean")

# estimate scaling so that we can pput the values in the same range.
scl = 3000/15
# add some labels

p = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum",  fill="cyan") + 
  geom_line(aes(x=year, y=tmax*scl), stat="summary", fun.y="mean", col="red") 
p 
p=p+ scale_y_continuous(sec.axis = sec_axis(~./scl, name=expression(paste("Maximum Temperature ",C**degree)))) + 
  labs(x="Year", y="Rainfall (mm/yr)")
p


# annotation. Put text on the plot.!! COOL.!!
p = ggplot(clim) + geom_bar(aes(x=year,y=rain), stat="summary", fun.y="sum",  fill="cyan") + 
  geom_line(aes(x=year, y=tmax*scl), stat="summary", fun.y="mean", col="red") 
p=p+ scale_y_continuous(sec.axis = sec_axis(~./scl, name=expression(paste("Maximum Temperature ",C**degree)))) + 
  labs(x="Year", y="Rainfall (mm/yr)")
p = p+ annotate("text", x=1990, y=500, label="Precip", colour="blue", size=6, hjust=0.5)
p = p+ annotate("text", x=1940, y=2700, label="Temperature", colour="red", size=6, hjust=0)
p

```

Now lets try *facets*  which allow you to separate things into different graphs
lets say we wanted a separate graph for each thinning intensity so  we could see how decade impacted recovery

```{r facet}
# first lets color by decade/scenario. Facet creates multiple graphs base on a dimension!!!Secen is the DECADE.
p2 = ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(scen)))+ stat_summary(fun.y="mean",geom="line", aes(col=as.factor(scen)))
p2
# now lets separate out the thinning intensities
p2= p2+facet_wrap(~as.factor(thin))
p2

# make prettier
p2 = ggplot(thindeep, aes(x=wy, y=plantc, col=as.factor(scen)))+ stat_summary(fun.y="mean",geom="line", size=2, aes(col=as.factor(scen)))
p2= p2+facet_wrap(~as.factor(thin)) 
p2 + scale_color_brewer(type="qual", palette="Set3", name="Decade") 
p2
```

